---
title: "Statistics and Quantitative Methods (S1)"
subtitle: "Week 7"
author: "Dr Stefano Coretta"
institute: "University of Edinburgh"
date: "2022/09/01"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css:
      - xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: "../../assets/macros.js"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, dpi = 300, out.height = "400px", fig.align = "center")
knitr::opts_knit$set(root.dir = here::here())
options(htmltools.dir.version = FALSE)
library(tidyverse)
theme_set(theme_light())
library(magrittr)
library(xaringanExtra)
use_xaringan_extra(c("panelset", "tachyons", "freezeframe"))
library(ggeffects)
library(marginaleffects)
options(show.signif.stars = FALSE)
library(broom.mixed)
library(sqmf)
data("shallow")
data("polite")
data("gestures")
data("dur_ita_pol")
dur_ita_pol %<>% drop_na(f1, v1_duration)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#23395b",
  text_font_google = google_font("Lato", "400", "700", "400i", "700i"),
  header_font_google = google_font("Montserrat", "400", "700", "400i", "700i"),
  code_font_google = google_font("Source Code Pro", "400", "700")
)
```

class: center middle inverse

# INTERACTIONS

---

layout: true

# Motivating interactions

---

Example 1: Shallow morphological parsing

```{r shallow}
shallow %>% select(ID, Group, Target, Relation_type, Branching, accuracy)
```

---

```{r shallow-2}
shallow <- shallow %>%
  filter(Relation_type != "NonConstituent") %>%
  mutate(
    Group = factor(Group, levels = c("L1", "L2")),
    Relation_type = factor(Relation_type, levels = c("Unrelated", "Constituent", "NonConstituent")),
    Branching = factor(Branching, levels = c("Left", "Right")),
    accuracy = factor(accuracy, levels = c("incorrect", "correct"))
  )
```

Let's filter the data to exclude `"NonConstituent"` from `Relation_type` and let's change character columns to factor columns.

---

.pull-left[

```{r shallow-plot-2, fig.show="hide"}
shallow %>%
  ggplot(aes(Relation_type, fill = accuracy)) +
  geom_bar(position = "fill") +
  facet_grid(~ Branching)
```

]

.pull-right[

![](`r knitr::fig_chunk('shallow-plot-2', 'png')`)

]

---

Let's try first with no interaction.

```{r shallow-lm}
shallow_lm <- glm(accuracy ~ Relation_type + Branching, data = shallow, family = binomial())
```

$$\text{accuracy}_{LO} = \beta_0 + \beta_1 \cdot \text{Relation_type}_{Const} + \beta_2 \cdot \text{Branching}_{Right}$$


<br>

--

$$\text{accuracy(Unrelated, Left)}_{LO} = \beta_0 + \beta_1 \cdot 0 + \beta_2 \cdot 0 = \beta_0$$

$$\text{accuracy(Constituent, Left)}_{LO} = \beta_0 + \beta_1 \cdot 1 + \beta_2 \cdot 0 = \beta_0 + \beta_1$$
$$\text{accuracy(Unrelated, Right)}_{LO} = \beta_0 + \beta_1 \cdot 0 + \beta_2 \cdot 1 = \beta_0 + \beta_2$$

$$\text{accuracy(Constituent, Right)}_{LO} = \beta_0 + \beta_1 \cdot 1 + \beta_2 \cdot 1 = \beta_0 + \beta_1 + \beta_2$$

---

```{r shallow-tidy}
# library(broom.mixed)
tidy(shallow_lm)
```

--

|                | Left          | Right                 |
|----------------|---------------|-----------------------|
| Unrelated      | 0.888         | 0.888 + 1.516         |
| Constituent    | 0.888 + 0.477 | 0.888 + 0.477 + 1.516 |

--

<br>

|                | Left  | Right |
|----------------|-------|-------|
| Unrelated      | 0.888 | 2.404 |
| Constituent    | 1.365 | 2.881 |

---

```{r shallow-preds}
ggpredict(shallow_lm, terms = c("Relation_type", "Branching"))
```

---

```{r shallow-preds-plot}
ggpredict(shallow_lm, terms = c("Branching", "Relation_type")) %>% plot()
```

---

```{r shallow-plot-2-2}
shallow %>%
  ggplot(aes(Relation_type, fill = accuracy)) +
  geom_bar(position = "fill") +
  facet_grid(~ Branching)
```

---

layout: false
layout: true

# Categorical + categorical: Example 1

---


Let's add an interaction term between `Relation_type` and `Branching`.

```{r shallow-lm-2}
shallow_lm_2 <- glm(
  accuracy ~ Relation_type + Branching + Relation_type:Branching,
  data = shallow, family = binomial()
)
```

$$\text{accuracy}_{LO} = \beta_0 + \beta_1 \cdot \text{Relation_type}_{Const} + \beta_2 \cdot \text{Branching}_{Right} + \beta_3 \cdot \text{Relation_type}_{Const} \cdot \text{Branching}_{Right}$$


<br>

--

$$\text{accuracy(Unrelated, Left)}_{LO} = \beta_0 + \beta_1 \cdot 0 + \beta_2 \cdot 0 + \beta_3 \cdot 1 \cdot 0 = \beta_0$$

$$\text{accuracy(Constituent, Left)}_{LO} = \beta_0 + \beta_1 \cdot 1 + \beta_2 \cdot 0 + \beta_3 \cdot 1 \cdot 0 = \beta_0 + \beta_1$$
$$\text{accuracy(Unrelated, Right)}_{LO} = \beta_0 + \beta_1 \cdot 0 + \beta_2 \cdot 1 + \beta_3 \cdot 0 \cdot 1 = \beta_0 + \beta_2$$

$$\text{accuracy(Constituent, Right)}_{LO} = \beta_0 + \beta_1 \cdot 1 + \beta_2 \cdot 1 + \beta_3 \cdot 1 \cdot 1 = \beta_0 + \beta_1 + \beta_2 + \beta_3$$

---

```{r shallow-lm-2-2}
tidy(shallow_lm_2)
```

--

|             | Left          | Right                          |
|-------------|---------------|--------------------------------|
| Unrelated   | 0.822         | 0.822 + 1.775                  |
| Constituent | 0.822 + 0.628 | 0.822 + 0.628 + 1.775 + -0.588 |

--

<br>

|             | Left  | Right |
|-------------|-------|-------|
| Unrelated   | 0.822 | 2.597 |
| Constituent | 1.45  | 2.637 |

---

```{r shallow-preds-2}
ggpredict(shallow_lm_2, terms = c("Relation_type", "Branching"))
```

---

```{r shallow-preds-plot-2}
ggpredict(shallow_lm_2, terms = c("Branching", "Relation_type")) %>% plot()
```

---

### Reporting the model and results

We fitted a linear model with a Bernoulli family to accuracy (incorrect vs correct) with the following predictors: relation type (unrelated vs constituent) and branching (left vs right).
We also included an interaction between relation type and branching.
Both predictors were coded with treatment contrasts and the first level listed above was set as the reference level.
We report in the following paragraph the results as probabilities (the model's estimates in log-odds are reported between parenthesis with standard errors).

According to the model, there is a 69% probability of obtaining a correct response when the relation type is unrelated and the word is left-branching ($\beta$ = 0.822, SE = 0.109).
When relation type is constituent and the word is left-branching, there probability increases to 81% ($\beta$ = 0.628, SE = 0.169).
When the word is right-branching and the relation type is unrelated, there is a 93% probability of obtaining a correct response ($\beta$ = 1.775, SE = 0.227).
The interaction term between relation type and branching indicates that when the word is right-branching and relation type is constituent there is a negligible difference in probability relative to relation type unrelated ($\beta$ = -0.588, SE = 0.331), so that in the right-branching condition relation type does not really affects the probability of getting a correct response.

---

layout: false
layout: true

# Categorical + categorical: Example 2

---

Example 2: Vowel duration and consonant voicing in Italian and Polish

- CVCV words where V1 = /a, o, u/ and C2 = /t, d, k, ɡ/.

- Investigate the effect of voicing of C2 (voiceless /t, k/, voiced /d, g/) on V1 duration.

- Data from Italian and Polish speakers.

---

```{r vdur}
dur_ita_pol %>% select(speaker, word, vowel, v1_duration, c2_phonation)
```


---

```{r vdur-voi-plot, warning=FALSE}
d_pos <- position_dodge(width = 0.8)

vdur_voi_plot <- ggplot(dur_ita_pol, aes(c2_phonation, v1_duration)) +
  geom_jitter(size = 1, alpha = 0.25, width = 0.2) +
  geom_violin(aes(fill = c2_phonation), width = 0.25) +
  stat_summary(fun = mean, position = d_pos, colour = "black") +
  labs(
    x = "Consonant voicing", y = "Vowel duration"
  ) +
  theme(legend.position = "none")
```

---

```{r vdur-voi-plot-2, out.height="500px", warning=FALSE, echo=FALSE}
vdur_voi_plot
```

---

```{r vdur-lang-plot, warning=FALSE}
vdur_lang_plot <- ggplot(dur_ita_pol, aes(language, v1_duration)) +
  geom_jitter(size = 1, alpha = 0.25, width = 0.2) +
  geom_violin(aes(fill = language), width = 0.25) +
  stat_summary(fun = mean, position = d_pos, colour = "black") +
  labs(
    x = "Language", y = "Vowel duration"
  ) +
  theme(legend.position = "none")
```

---

```{r vdur-lang-plot-2, out.height="500px", echo=FALSE, warning=FALSE}
vdur_lang_plot
```

---

```{r vdur-voi-lang-plot, warning=FALSE}
jd_pos <- position_jitterdodge(dodge.width = 0.8, jitter.width = 0.3)
d_pos <- position_dodge(width = 0.8)

vdur_plot <- ggplot(dur_ita_pol, aes(language, v1_duration)) +
  geom_jitter(aes(colour = c2_phonation), size = 1, alpha = 0.25, position = jd_pos) +
  geom_violin(aes(fill = c2_phonation), position = d_pos, width = 0.5) +
  stat_summary(aes(group = paste(c2_phonation, language)), fun = mean, position = d_pos, colour = "black") +
  labs(
    x = "Language", y = "Vowel duration"
  ) +
  theme(legend.position = "none")
```

---

```{r vdur-voi-lang-plot-2, warning=FALSE, echo=FALSE, out.height="500px"}
vdur_plot
```


???

Is the effect of consonant voicing on vowel duration the same in Italian and Polish?

To answer this question we need to include an interaction between consonant voicing and language.

---

```{r vdur-lm}
vdur_lm <- lm(v1_duration ~ c2_phonation * language, data = dur_ita_pol)
# NOTE: "c2_phonation * language" is a short-cut for
# "c2_phonation + language + c2_phonation:language"
```

$$\text{v1_duration} = \beta_0 + \beta_1 \cdot \text{c2_phonation} + \beta_2 \cdot \text{language} + \beta_3 \cdot \text{c2_phonation} \cdot \text{language}$$

<br>

--

$$\text{v1_duration(voiced, Italian)} = \beta_0 + \beta_1 \cdot 0 + \beta_2 \cdot 0 + \beta_3 \cdot 0 \cdot 0= \beta_0$$

$$\text{v1_duration(voiceless, Italian)} = \beta_0 + \beta_1 \cdot 1 + \beta_2 \cdot 0 + \beta_3 \cdot 1 \cdot 0 = \beta_0 + \beta_1$$

$$\text{v1_duration(voiced, Polish)} = \beta_0 + \beta_1 \cdot 0 + \beta_2 \cdot 1 + \beta_3 \cdot 0 \cdot 1 = \beta_0 + \beta_2$$

$$\text{v1_duration(voiceless, Polish)} = \beta_0 + \beta_1 \cdot 1 + \beta_2 \cdot 1 + \beta_3 \cdot 1 \cdot 1 = \beta_0 + \beta_1 + \beta_2 + \beta_3$$

---

```{r}
tidy(vdur_lm)
```

---

```{r vdur-lm-pred}
ggpredict(vdur_lm, terms = c("c2_phonation", "language"))
```

---

```{r vdur-lm-pred-plot}
ggpredict(vdur_lm, terms = c("c2_phonation", "language")) %>%
  plot()
```

???


```{r}
marginaleffects(vdur_lm, variables = "c2_phonation", by = "language")
```


---

layout: false
layout: true

# Categorical + continous

---

```{r dur-ita}
dur_ita <- dur_ita_pol %>%
  filter(language == "Italian") %>%
  mutate(
    vowel = factor(vowel, levels = c("a", "o", "u"))
  )

dur_ita %>% select(word, vowel, v1_duration, c2_duration)
```

---

```{r dur-ita-plot}
dur_ita %>%
  ggplot(aes(c2_duration, v1_duration)) +
  geom_point(alpha = 0.25) +
  geom_smooth(aes(colour = vowel), method = "lm")
```

???

There seems to be a positive effect of consonant duration on vowel duration: i.e. the longer the consonant the longer the vowel.

But is the effect of consonant duration on vowel duration the same for the three vowels /a, o, u/?

---

```{r vdur-lm-2}
vdur_lm_2 <- lm(v1_duration ~ c2_duration * vowel, data = dur_ita)
```

$$
\begin{align}
\text{v1_duration} = & \beta_0 + \beta_1 \cdot \text{c2_duration} + \beta_2 \cdot \text{vowel_o} + \beta_3 \cdot + \text{vowel_u} \\
& + \beta_4 \cdot \text{c2_duration} \cdot \text{vowel_o} + \beta_5 \cdot \text{c2_duration} \cdot \text{vowel_u}
\end{align}
$$

<br>

--

$$\text{v1_duration(c2_dur = 0, a)} = \beta_0 + \beta_1 \cdot 0 + \beta_2 \cdot 0 + \beta_3 \cdot 0 \cdot 0= \beta_0$$

$$\text{v1_duration(c2_dur = 0, o)} = \beta_0 + \beta_1 \cdot 0 + \beta_2 \cdot 1 + \beta_3 \cdot 0 + \beta_4 \cdot 0 \cdot 1 + \beta_5 \cdot 0 \cdot 0 = \beta_0 + \beta_2$$

$$\text{v1_duration(c2_dur = 0, u)} = \beta_0 + \beta_1 \cdot 0 + \beta_2 \cdot 0 + \beta_3 \cdot 1 + \beta_4 \cdot 0 \cdot 0 + \beta_5 \cdot 0 \cdot 1 = \beta_0 + \beta_3$$

---

$$
\begin{align}
\text{v1_duration} = & \beta_0 + \beta_1 \cdot \text{c2_duration} + \beta_2 \cdot \text{vowel_o} + \beta_3 \cdot + \text{vowel_u} \\
& + \beta_4 \cdot \text{c2_duration} \cdot \text{vowel_o} + \beta_5 \cdot \text{c2_duration} \cdot \text{vowel_u}
\end{align}
$$

<br>

--

$$\text{v1_duration(c2_dur = 1, a)} = \beta_0 + \beta_1 \cdot 1 + \beta_2 \cdot 0 + \beta_3 \cdot 0 \cdot 0= \beta_0 + \beta_1$$

$$\text{v1_duration(c2_dur = 1, o)} = \beta_0 + \beta_1 \cdot 1 + \beta_2 \cdot 1 + \beta_3 \cdot 0 + \beta_4 \cdot 1 \cdot 1 + \beta_5 \cdot 1 \cdot 0 = \beta_0 + \beta_1 + \beta_2 + \beta_4$$

$$\text{v1_duration(c2_dur = 1, u)} = \beta_0 + \beta_1 \cdot 1 + \beta_2 \cdot 0 + \beta_3 \cdot 1 + \beta_4 \cdot 1 \cdot 0 + \beta_5 \cdot 1 \cdot 1 = \beta_0 + \beta_1 + \beta_3 + \beta_5$$

---

```{r vdur-lm-2-tidy}
tidy(vdur_lm_2)
```

---

.pull-left[
```{r vdur-lm-2-pred, eval=FALSE}
ggpredict(
  vdur_lm_2, 
  terms = c(
    "c2_duration [0, 50, 100]",
    "vowel"
  )
)
```
]

.pull-right[
```{r vdur-lm-2-pre-2d, echo=FALSE}
ggpredict(vdur_lm_2, terms = c("c2_duration [0, 50, 100]", "vowel"))
```
]

???

You can specify specific values for continuous predictors, like in `"c2_duration [0, 50, 100]"`.

Here, I am asking `ggpredict()` to return predicted values for when `c2_duration` is 0, 50 and 100 ms.

To learn more about how to specify values, check the documentation in `?ggpredict`.

---

```{r vdur-lm-2-pred-plot}
ggpredict(vdur_lm_2, terms = c("c2_duration", "vowel")) %>%
  plot()
```

???

Note that for plotting you should not give specific values for the continuous predictor that goes into the *x*-axis, because this axis will include the entire range of values for that continuous predictor.

---

layout: false
layout: true

# Continuous + continuous

---

```{r vdur-rate}
dur_ita %>%
  ggplot(aes(c2_duration, v1_duration, colour = speech_rate)) +
  geom_point(alpha = 0.5)
```

???

Let's disregard vowel for the time being (you will add vowel in in the tutorial).

---

```{r vdur-lm-3}
vdur_lm_3 <- lm(v1_duration ~ c2_duration * speech_rate, data = dur_ita)

tidy(vdur_lm_3)
```

---

```{r vdur-lm-3-plot-1}
ggpredict(vdur_lm_3, terms = c("c2_duration", "speech_rate [3, 5, 7]")) %>%
  plot()
```

---

```{r vdur-lm-3-plot-2}
ggpredict(vdur_lm_3, terms = c("speech_rate", "c2_duration [50, 100, 150]")) %>%
  plot()
```

