---
title: "Statistics and Quantitative Methods (S1)"
subtitle: "Week 9 — Workshop"
author: "Dr Stefano Coretta"
institute: "University of Edinburgh"
date: "2022/11/15"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css:
      - xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: "../../assets/macros.js"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, dpi = 300, fig.width = 7, fig.height = 5, out.height = "500px", fig.align = "center")
knitr::opts_knit$set(root.dir = here::here())
options(htmltools.dir.version = FALSE)

library(xaringanExtra)
use_xaringan_extra(c("panelset", "tachyons", "freezeframe"))
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#23395b",
  text_font_google = google_font("Lato", "400", "700", "400i", "700i"),
  header_font_google = google_font("Montserrat", "400", "700", "400i", "700i"),
  code_font_google = google_font("Source Code Pro", "400", "700"),
  link_color = "#d95f02"
)
```


# SQMF workshop attendance

**THERE IS A NEW SYSTEM IN PLACE**

You need to record your attendance by scanning the QR code outside of the room.

---

# R MARKDOWN TUTORIAL

- There is no new tutorial file for this week.

- You can keep working on previous tutorials if you haven't completed them.

- If you have completed the previous tutorials, you can take a look at the *p*-values of the linear models you fitted so far.

  - Remember to attach the lmerTest package to get *p*-values for `g/lmer()` models.
