---
title: "Statistics and Quantitative Methods (S1)"
subtitle: "Week 8"
author: "Dr Stefano Coretta"
institute: "University of Edinburgh"
date: "2022/11/08"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css:
      - xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: "../../assets/macros.js"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, dpi = 300, out.height = "400px", fig.align = "center")
knitr::opts_knit$set(root.dir = here::here())
options(htmltools.dir.version = FALSE)
library(tidyverse)
theme_set(theme_light())
library(magrittr)
library(xaringanExtra)
use_xaringan_extra(c("panelset", "tachyons", "freezeframe"))
library(ggeffects)
library(marginaleffects)
options(show.signif.stars = FALSE)
library(broom.mixed)
library(sqmf)
data("mald_1_1")
data("shallow")
data("polite")
data("gestures")
data("dur_ita_pol")
dur_ita_pol %<>% drop_na(f1, v1_duration)
library(lme4)
library(patchwork)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#23395b",
  text_font_google = google_font("Lato", "400", "700", "400i", "700i"),
  header_font_google = google_font("Montserrat", "400", "700", "400i", "700i"),
  code_font_google = google_font("Source Code Pro", "400", "700")
)
```

class: center middle

.f1.link.dim.br3.ph3.pv2.mb2.dib.white.bg-purple[
[.white[Linear models illustrated]](https://stefanocoretta.shinyapps.io/lines/)
]

---

layout: true

## Repeated measures: Multiple subjects

---

```{r mald}
mald_1_1
```

---

```{r mald-acc}
mald_1_1 %>%
  mutate(ACC_num = ifelse(ACC == "correct", 1, 0)) %>%
  ggplot(aes(PhonLev, ACC_num, colour = IsWord)) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial"))
```

---

layout: false

class: center middle

.f1.link.dim.br3.ph3.pv2.mb2.dib.white.bg-purple[
[.white[Linear models illustrated]](https://stefanocoretta.shinyapps.io/lines/)
]

---

layout: true

## Random effects: varying intercept

---

```{r mald-lm}
mald_lm <- glm(ACC ~ IsWord * PhonLev, data = mald_1_1, family = binomial())

tidy(mald_lm)
```

---

```{r mald-lm-p}
mald_lm_p <- ggpredict(mald_lm, terms = c("PhonLev [all]", "IsWord")) %>% plot(limits = c(0.75, 1))
mald_lm_p
```

---

```{r mald-subj}
subjs <- sample(mald_1_1$Subject, 9)
mald_1_1 %>%
  filter(Subject %in% subjs) %>%
  ggplot(aes(Subject, fill = ACC)) +
  geom_bar(position = "fill")
```

---

```{r mald-lm-2}
library(lme4)
mald_lm_2 <- glmer(ACC ~ IsWord * PhonLev + (1 | Subject), data = mald_1_1, family = binomial())

summary(mald_lm_2)
```

---

```{r mald-lm-2-p}
mald_lm_2_p <- ggpredict(mald_lm_2, terms = c("PhonLev [all]", "IsWord")) %>% plot(limits = c(0.75, 1))
mald_lm_2_p
```

---

```{r mald-lm-lm-p}
# library(patchwork)
mald_lm_p + mald_lm_2_p + plot_layout(nrow = 2, guides = "collect")
```

---

Let's check the **conditional means**

```{r mald-lm-2-r}
mald_lm_2_r <- ranef(mald_lm_2) %>% as_tibble()
mald_lm_2_r
```

---

```{r mald-lm-2-r-p}
mald_lm_2_r %>%
  ggplot(aes(condval, grp)) +
  geom_vline(xintercept = 0, colour = "#7fc97f") +
  geom_pointrange(
    aes(xmin = condval + 1.96 * condsd, xmax = condval - 1.96 * condsd),
    alpha = 0.7
  )
```

---

```{r mald-lm-2-pred-r}
ggpredict(mald_lm_2, terms = c("PhonLev [all]", "IsWord", "Subject [sample=8]"), type = "re") %>% plot()
```


---

layout: false

class: center middle

.f1.link.dim.br3.ph3.pv2.mb2.dib.white.bg-purple[
[.white[Linear models illustrated]](https://stefanocoretta.shinyapps.io/lines/)
]

---

layout: true

## Random effects: varying slopes

---

```{r mald-subj-2, warning=FALSE}
mald_1_1 %>%
  filter(Subject %in% sample(mald_1_1$Subject, 9)) %>%
  mutate(ACC_num = ifelse(ACC == "correct", 1, 0)) %>%
  ggplot(aes(PhonLev, ACC_num, colour = IsWord)) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  facet_wrap(~ Subject)
```

---

```{r mald-lm-3}
mald_lm_3 <- glmer(ACC ~ IsWord * PhonLev + (IsWord | Subject), data = mald_1_1, family = binomial())

summary(mald_lm_3)
```

---

```{r mald-lm-3-p}
mald_lm_3_p <- ggpredict(mald_lm_3, terms = c("PhonLev [all]", "IsWord")) %>% plot(limits = c(0.75, 1))
mald_lm_3_p
```

---

```{r mald-lm-lm-3-p}
mald_lm_p + mald_lm_3_p + plot_layout(nrow = 2, guides = "collect")
```

---

Let's check the **conditional means**

```{r mald-lm-3-r}
mald_lm_3_r <- ranef(mald_lm_3) %>% as_tibble()
mald_lm_3_r
```

---

```{r mald-lm-3-r-p}
mald_lm_3_r %>%
  filter(term == "IsWordFALSE") %>%
  ggplot(aes(condval, reorder(grp, condval))) +
  geom_vline(xintercept = 0, colour = "#7fc97f") +
  geom_pointrange(
    aes(xmin = condval + 1.96 * condsd, xmax = condval - 1.96 * condsd),
    alpha = 0.7
  )
```

---

```{r mald-lm-3-pred-r}
ggpredict(mald_lm_3, terms = c("PhonLev [all]", "IsWord", "Subject [sample=9]"), type = "re") %>% plot()
```

---

layout: false
layout: true

## Repeated measures: Multiple subjects and words

---

```{r shallow}
shallow_filt <- shallow %>%
  filter(Branching == "Left")

shallow_filt
```

---

```{r id-target}
unique(shallow_filt$ID)

unique(shallow_filt$Target)
```

--

`ID` and `Target` are crossed, i.e. each subject has seen all the targets and each target has been seen by each subject.

We talk about **crossed random effects**.

---

```{r shal-lm}
shal_lm <- lmer(
  RT ~
    Relation_type * Group +
    (1 | ID) + (1 | Target),
  data = shallow_filt
)

summary(shal_lm)
```

---

```{r shal-lm-p}
shal_lm_p <- ggpredict(shal_lm, terms = c("Group", "Relation_type")) %>% plot()
shal_lm_p
```

---

```{r shal-lm-r}
shal_lm_r <- ranef(shal_lm) %>% as_tibble()
shal_lm_r
```

---

```{r shal-lm-r-p}
shal_lm_r %>%
  filter(grpvar == "ID") %>%
  ggplot(aes(condval, reorder(grp, condval))) +
  geom_vline(xintercept = 0, colour = "#7fc97f") +
  geom_pointrange(
    aes(xmin = condval + 1.96 * condsd, xmax = condval - 1.96 * condsd),
    alpha = 0.7
  )
```

---

```{r shal-lm-r-p-2}
shal_lm_r %>%
  filter(grpvar == "Target") %>%
  ggplot(aes(condval, reorder(grp, condval))) +
  geom_vline(xintercept = 0, colour = "#7fc97f") +
  geom_pointrange(
    aes(xmin = condval + 1.96 * condsd, xmax = condval - 1.96 * condsd),
    alpha = 0.7
  )
```

---

```{r shal-lm-pred-r-id}
ggpredict(shal_lm, terms = c("ID [sample=6]", "Relation_type"), type = "re") %>% plot()
```

---

```{r shal-lm-pred-r-target}
ggpredict(shal_lm, terms = c("Target [sample=6]", "Relation_type"), type = "re") %>% plot()
```

---

```{r shal-lm-2}
shal_lm_2 <- lmer(
  RT ~
    Relation_type * Group +
    (Relation_type | ID) + (Relation_type | Target),
  data = shallow_filt
)

summary(shal_lm_2)
```

---

```{r shal-lm-2-p}
shal_lm_2_p <- ggpredict(shal_lm_2, terms = c("Group", "Relation_type")) %>% plot()
shal_lm_2_p
```


---

```{r shal-lm-2-pred-r-id}
ggpredict(shal_lm_2, terms = c("ID [sample=10]", "Relation_type"), type = "re") %>% plot()
```

---

```{r shal-lm-2-pred-r-target}
ggpredict(shal_lm_2, terms = c("Target [sample=6]", "Relation_type"), type = "re") %>% plot()
```

