---
title: "Statistics and Quantitative Methods (S1)"
subtitle: "Week 10"
author: "Dr Stefano Coretta"
institute: "University of Edinburgh"
date: "2022/11/23"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css:
      - xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: "../../assets/macros.js"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, dpi = 300, out.height = "500px", fig.align = "center")
knitr::opts_knit$set(root.dir = here::here())
options(htmltools.dir.version = FALSE)
library(tidyverse)
theme_set(theme_light())
library(magrittr)
library(xaringanExtra)
use_xaringan_extra(c("panelset", "tachyons", "freezeframe"))
library(ggeffects)
library(marginaleffects)
options(show.signif.stars = FALSE)
library(broom.mixed)
library(sqmf)
data("mald_1_1")
data("shallow")
data("polite")
data("gestures")
data("dur_ita_pol")
dur_ita_pol %<>% drop_na(f1, v1_duration)
library(lme4)
library(patchwork)
options(ggplot2.discrete.fill = RColorBrewer::brewer.pal(12, "Dark2"))
options(ggplot2.discrete.colour = RColorBrewer::brewer.pal(12, "Dark2"))
```


```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#23395b",
  text_font_google = google_font("Lato", "400", "700", "400i", "700i"),
  header_font_google = google_font("Montserrat", "400", "700", "400i", "700i"),
  code_font_google = google_font("Source Code Pro", "400", "700"),
  base_font_size = "18px"
)
```

layout: true

## Example of plotting

---

.pull-left[
One continuous variable: density

```{r cont-dens, fig.show='hide'}
dur_ita_pol %>%
  ggplot(aes(v1_duration)) +
  geom_density(fill = "gray")
```
]

.pull-right[
![:scale 90%](`r knitr::fig_chunk('cont-dens', 'png')`)
]

---

.pull-left[
One continuous variable: density (with categorical variable)

```{r cont-dens-1, fig.show='hide'}
dur_ita_pol %>%
  ggplot(
    aes(v1_duration, fill = c2_phonation)
  ) +
  geom_density(alpha = 0.5)
```
]

.pull-right[
![:scale 90%](`r knitr::fig_chunk('cont-dens-1', 'png')`)
]

---

.pull-left[
Two continuous variables

```{r cont-cont, fig.show='hide'}
dur_ita_pol %>%
  ggplot(aes(speech_rate_c, v1_duration)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm")
```
]

.pull-right[
![:scale 90%](`r knitr::fig_chunk('cont-cont', 'png')`)
]

---

.pull-left[
Two continuous variables and one categorical variable

```{r cont-cont-cat, fig.show='hide'}
dur_ita_pol %>%
  ggplot(
    aes(speech_rate_c, v1_duration, colour = vowel)
  ) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm")
```
]

.pull-right[
![:scale 90%](`r knitr::fig_chunk('cont-cont-cat', 'png')`)
]

---

.pull-left[
Two continuous variables and one categorical variable

```{r cont-cont-cat-1, fig.show='hide'}
dur_ita_pol %>%
  ggplot(aes(speech_rate_c, v1_duration)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm",
              aes(colour = vowel)) +
  facet_grid(~ vowel)
```
]

.pull-right[
![:scale 90%](`r knitr::fig_chunk('cont-cont-cat-1', 'png')`)
]

---

.pull-left[
One continuous variable and one categorical variable

```{r cont-cat, fig.show='hide'}
polite %>%
  ggplot(aes(attitude, f0mn)) +
  geom_jitter(width = 0.2)
```
]

.pull-right[
![:scale 90%](`r knitr::fig_chunk('cont-cat', 'png')`)
]

---

.pull-left[
One continuous variable and two categorical variables

```{r cont-cat-cat, fig.show='hide'}
polite %>%
  ggplot(aes(attitude, f0mn,
             colour = musicstudent)) +
  geom_point(
    position = position_jitterdodge(
      jitter.width = 0.2,
      dodge.width = 0.8
    )
  )
```
]

.pull-right[
![:scale 90%](`r knitr::fig_chunk('cont-cat-cat', 'png')`)
]

---

.pull-left[
One continuous variable and three categorical variables

```{r cont-cat-cat-cat, fig.show='hide'}
polite %>%
  ggplot(aes(attitude, f0mn,
             colour = musicstudent)) +
  geom_point(
    position = position_jitterdodge(
      jitter.width = 0.2,
      dodge.width = 0.8)
  ) +
  facet_grid(~ gender)
```
]

.pull-right[
![:scale 90%](`r knitr::fig_chunk('cont-cat-cat-cat', 'png')`)
]

---


.pull-left[
One continuous variable and three categorical variables

```{r cont-cat-cat-cat-1, fig.show='hide'}
polite %>%
  ggplot(aes(attitude, f0mn,
             fill = musicstudent)) +
  geom_violin() +
  geom_point(
    position = position_jitterdodge(
      jitter.width = 0.05,
      dodge.width = 0.85
    ),
    alpha = 0.5
  ) +
  facet_grid(~ gender)
```
]

.pull-right[
![:scale 90%](`r knitr::fig_chunk('cont-cat-cat-cat-1', 'png')`)
]

---

.pull-left[
One continuous variable and four categorical variables

```{r cont-cat-4, fig.show='hide'}
dur_ita_pol %>%
  ggplot(aes(c2_place, c2_clos_duration,
             fill = c2_phonation)) +
  geom_violin() +
  geom_point(
    position = position_jitterdodge(
      jitter.width = 0.05,
      dodge.width = 0.9
    ),
    alpha = 0.25
  ) +
  facet_grid(language ~ vowel)
```
]

.pull-right[
![:scale 90%](`r knitr::fig_chunk('cont-cat-4', 'png')`)
]

---

.pull-left[
One categorical variable

```{r cat, fig.show='hide'}
shallow %>%
  ggplot(aes(accuracy)) +
  geom_bar()
```
]

.pull-right[
![:scale 90%](`r knitr::fig_chunk('cat', 'png')`)
]

---

.pull-left[
Two categorical variables

```{r cat-cat, fig.show='hide'}
shallow %>%
  ggplot(aes(Relation_type,
             fill = accuracy)) +
  geom_bar(position = "fill")
```
]

.pull-right[
![:scale 90%](`r knitr::fig_chunk('cat-cat', 'png')`)
]

---

layout: false
layout: true

## Random effects
